Optimal Fitting of Planar Curves to Prescribed Constraints

### 2013-08-08 idp/documentation/paper.tex
replaced template bibliography entry
added some comments for the included packages, added hyperref package to enable pdf index, continued working on section about spiro splines

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{a4wide}
\usepackage{parskip}
\usepackage{times}
\usepackage{textcomp}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{multirow}

\clubpenalty = 10000
\widowpenalty = 10000

@@

% TODO: consider splitting some illustrations into multiple figures if there is still space left at the end

@@

		% TODO: make sure introduction does not overlap with start of next section regarding both content and wording

@@

			At this point, it seems like the two ways in which curve design software can affect the design process are through the choice of language and fairness measure. Ideally, the language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently. The fairness should capture the intuitive notion of curve smoothness sufficiently well. These goals are competing with the software's ability to derive curves from descriptions and the fairness measure reliably and efficiently, a process which becomes increasingly difficult with more expressive description languages and more complex fairness measures. From these considerations, we think that applying the following criteria is a good start for assessing the usability of curve design tools.

@@

				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing curvature continuity. It is difficult for humans to choose the coefficient points in such a way that curvature continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard. The problem is aggravated by the fact that curvature incontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes. In case the software allows the user to request curvature-continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result. Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with language expressiveness.

@@

				% TODO: consider removing the handle markers to make curvature incontinuity more visible
				Example 4 shows a spline with a curvature incontinuity which went unnoticed by the artist at the time of creation. In fact, this incontinuity would be even less visible if the curve were not vertical at said point, manifesting itself only in a reduced overall smoothness of the curve, without a clear indication of where the problem lies.

@@

@@

		% demonstrate euler spirals and bézier curves being embedded
		% revisit examples made when analyzing Bézier splines and Spiro splines
		% a bézier node comes with 2 handles, making it 3 points, or 6 scalar values that are specified, compared to 2 scalars for a DC specification

@@

		\bibitem{CranePatterson2000}
			\emph{History of the mind-body problem}\\
			Tim Crane, Sarah Patterson\\
			Routledge, 2000

>>>>

\usepackage[utf8]{inputenc}    % enables use of utf8 characters in the input file
\usepackage[T1]{fontenc}       % enables proper output of unicode characters such that text can be copy-pasted from the pdf file
\usepackage{hyperref}          % enables pdf index, links for web addresses, intra-document links
\usepackage{a4wide}            % reduces page margins
\usepackage{parskip}           % replaces hanging paragraphs with spacing between paragraphs
\usepackage{times}             % replaces Computer Modern font with something resembling Times
\usepackage{textcomp}          % adds symbols like \textdegree
\usepackage{mathtools}         % modern version of ams math package
\usepackage{amssymb}           % ams math symbols
\usepackage{multirow}          % enables multirow tables

% penalizes lonely lines at the beginning or end of a page
\clubpenalty = 10000
\widowpenalty = 10000

@@

% TODO: consider splitting some illustrations into multiple figures if there is still space left at the end
% TODO: replace curvature continuity with G^2 continuity

@@

		% TODO: make sure introduction does not overlap with start of next section regarding both content and wording
		% TODO: comment on that although there has been some research into better curves (quote MVC curves), it was less focused on usability and hasn't had much impact on actual vector graphics software which continues to almost exclusively use bézier splinse to this day

@@

			We conclude that the two ways in which curve design software can affect the design process are through the choice of language and fairness measure. Ideally, the language should be very expressive, yet easy to handle for humans, enabling the user to communicate their intent to the software efficiently. The fairness should capture the intuitive notion of curve smoothness sufficiently well. These goals are competing with the software's ability to derive curves from descriptions and the fairness measure reliably and efficiently, a process which becomes increasingly difficult with more expressive description languages and more complex fairness measures. From these considerations, we think that applying the following criteria is a good start for assessing the usability of curve design tools.

@@

				The issue with the fairness measure is that it always selects the uniquely determined cubic polynomial, thus not guaranteeing curvature continuity. It is difficult for humans to choose the coefficient points in such a way that curvature continuity is established, making it hard to create smooth curves if the design software does not support the user in this regard. The problem is aggravated by the fact that curvature discontinuities can be hard to spot and may thus only be discovered much later in the design process, increasing the cost of these mistakes. In case the software allows the user to request curvature-continuous Bézier splines, this sacrifices some degree of freedom for each node, thus restricting the way nodes can be manipulated and in turn possibly forcing the user to place more nodes in order to get the desired result. Placing more nodes leads to some disadvantages that will be discussed further in the paragraphs dealing with language expressiveness.

@@

				% TODO: consider removing the handle markers to make curvature discontinuity more visible
				Example 4 shows a spline with a curvature discontinuity which went unnoticed by the artist at the time of creation. In fact, this discontinuity would be even less visible if the curve were not vertical at said point, manifesting itself only in a reduced overall smoothness of the curve, without a clear indication of where the problem lies.

@@

				Spiro splines were added to Inkscape in 2009. Even though they were integrated using path effects and thus have a somewhat cumbersome user interface, they constitute a commendable effort of incorporating results from curve research into vector graphics software, something that has hardly happened this far. More information on Spiro splines can be found in the dissertation of Raphael Levien \cite{Levien2009}.

				Spiro splines consist of parts of the Euler spiral, which is a curve whose curvature changes linearly with the curve's arc length. The user can specify control points through which the software puts an interpolating spline. There is no way to directly specify tangent direction and/or curvature. The splines are guaranteed to have \(G^2\) continuity.

				Again, the model from section \ref{section:curve_design_process} is applied. With Spiro splines, the modelling is fairly straightforward. The user supplies a set of points for which the software computes an interpolating spline. The fairness measure chooses the segments of this spline to be parts of the Euler spiral and makes sure that there is \(G^2\) continuity and no excessive winding.

				We analyze Spiro splines in terms of the criteria presented in section \ref{section:usability_curve_design_tools}. The language of Spiro splines being simply points on the curve, plus the fact that on sufficiently smooth curves, not very many of them being needed, the language is well-suited to be read and spoken by humans (criteria 2 and 3). The fairness measure leads to beautifully smooth curves that are ususally sufficiently close to what a human would expect given the control points (criterion 4). The software is also very good and sufficiently quick at constructing these curves (there are some cases where the algorithm fails to find a meaningful curve, but these cases are fairly artificial and can usually be avoided when designing actual curves) and thus fulfills criterion 5 well enough.

				The only real issue lies in the expressiveness of the language.

@@

		% demonstrate euler spirals and bézier curves being embedded
		%   minimization of normal jerk squares prefers linear change in curvature
		% revisit examples made when analyzing Bézier splines and Spiro splines
		% a bézier node comes with 2 handles, making it 3 points, or 6 scalar values that are specified, compared to 2 scalars for a DC specification

@@

		\bibitem{Levien2009}
			\emph{From Spiral to Spline: Optimal Techniques in Interactive Curve Design}\\
			Raphael Linus Levien\\
			University of California, Berkeley, Fall 2009
